Pretty darn good control: when are approximate solutions better than approximate models

### 2023-06-29 manuscript/manuscript.Rmd
started future directions

Our findings paint a picture of how a single-species optimal management strategy may loose performance rather dramatically when controlling a more complex ecosystem.
Here, DRL performs better from both economical *and* conservation points of view.
Moreover, rather unintuitively, CMort -- known to be suboptimal and unsustainable for single-species models -- can turn out to even outperform the CEsc -- the single-species optimal strategy -- for complex ecosystems.
Finally, within this regime of complex, possibly varying, ecosystems, we show that DRL consistently finds a policy which effectively either matches the best classical strategy, or outperforms it.
We strengthen this result with a stability investigation: we show that random perturbations in the model's parameter values used do not significantly vary the conclusion that DRL outperforms constant escapement.

@@

## Pasted to-be-discussed

Eqs. \eqref{eq:3d model} are a generalization of the single-species model \eqref{eq:may}: this can be seen by setting 
\begin{align*}
  c_{XY} = d = f = \sigma_Z = 0.
\end{align*}
Throughout most of our paper, we use the parameter values given in eq. \eqref{eq:param vals}.
In particular, $c_{XY} = d = f = 0.1$ and $\sigma_Z=0.05$ so, in broad strokes, our three-species model is a slight perturbation of the single-species model.

Here we note two surprising things:
First, optimal constant escapement performs considerably worse than not only DRL-based methods, but also than MSY.
This conceptually shows how one-dimensional intuitions on ecological control may loose validity for complex ecosystems. **(this might be unwelcome by the reviewers if not discussed properly)**
Second, we see that classical strategies have relatively spread out reward distributions as compared to DRL methods.

@@

Because of the simplicity of the classical policy functions, the optimal such policy may be easily estimated through a grid-search.
This is the case since these policy functions are only specified by one or two parameters (respectively in the single fishery, or two fishery cases).
In contrast, DRL optimizes over the more expressive -- and more complicated -- family of policies parametrized by a neural network.
Because of this, our DRL approach uses a variant of *gradient ascent* (the proximal policy optimization algorithm) to obtain a good candidate policy.

We showed that for sufficiently complex management scenarios -- with multiple harvested species and, possibly, time-varying parameters -- DRL-based management strategies perform significantly better than constant escapement.
This with respect to both average rewards received and conservation goals.
In this sense, an approximate solution to a more complicated and expressive model, can outperform the optimal solution of the single-species problem---even when the parameters of the single-species solution are empirically optimized.

What is more, this DRL advantage happened within a regime of parameter space where species interaction terms were much smaller than the single-species logistic growth terms.
Indeed, interaction parameters in \eqref{eq:3d model} are about an order of magnitude smaller than $r_X$ and $r_Y$.
In this situation logistic growth is a reasonable approximation of the system, and thus a constant escapement solution is a plausible control strategy.

We found that the optimal constant mortality policy surprisingly performs much better than constant escapement (Fig. \ref{fig:main}).
However, it can be observed in Fig. \ref{fig:fracmsy} that the constant mortality strategy faces a trade-off: high sustainability is achieved for sub-optimal constant mortality rates, but only at a significant decrease in the mean episode reward.
We expect that with increasing ecosystem complexity this phenomenon might become more pronounced.
We can understand this as a consequence of the rigidity of classical strategies: the simplicity of their expressions, depending only on a few parameters, means that policy optimization is constrained to a rather reduced subset of the space of possible policies.

@@

@@

# Appendix

## Results for stationary models

@@

## PPO policy function for non-stationary model

@@

## Gaussian process interpolation

>>>>

Our findings paint a picture of how a single-species optimal management strategy may lose performance rather dramatically when controlling a more complex ecosystem.
Here, DRL performs better from both economical *and* conservation points of view.
Moreover, rather unintuitively, CMort -- known to be suboptimal and unsustainable for single-species models -- can turn out to even outperform the CEsc -- the single-species optimal strategy -- for complex ecosystems.
Finally, within this regime of complex, possibly varying, ecosystems, we show that DRL consistently finds a policy which effectively either matches the best classical strategy, or outperforms it.
We strengthen this result with a stability investigation: we show that random perturbations in the model's parameter values used do not significantly vary the conclusion that DRL outperforms constant escapement.

@@

@@

Because of the simplicity of the classical policy functions, the optimal such policy may be easily estimated through a grid-search.
This is the case since these policy functions are only specified by one or two parameters (respectively in the single fishery, or two fishery cases).
In contrast, DRL optimizes over the more expressive -- and more complicated -- family of policies parametrized by a neural network.
Neural networks are often used as flexible function approximators that can be efficiently optimized over.

We showed that for sufficiently complex management scenarios -- Models 3 and 4 -- DRL-based management strategies perform significantly better than CEsc.
This with respect to both average rewards received and conservation goals.
In this sense, an approximate solution to a more complicated and expressive model, can outperform the optimal solution of the single-species problem---even when the parameters of the single-species solution are empirically optimized.

We found that the optimal CMort policy surprisingly performs much better than constant escapement (Fig. \ref{fig:main}).
However, it can be observed in Fig. \ref{fig:fracmsy} that the constant mortality strategy faces a trade-off: high sustainability is achieved for sub-optimal constant mortality rates, but only at a significant decrease in the mean episode reward.
We expect that with increasing ecosystem complexity this phenomenon might become more pronounced.
We can understand this as a consequence of the rigidity of classical strategies: the simplicity of their expressions, depending only on a few parameters, means that policy optimization is constrained to a rather reduced subset of the space of possible policies.

The question of when multi-species models are well-approximated by single-species models was studied in detail in [@burgess].
Here our approach is dual to that of the aforementioned paper.
Rather than first optimizing a single-species model to approximate a more complex model and then finding the MSY value for the single-species model, we used simulated data to optimize CMort and CEsc directly on the three-species model.
We do not investigate further whether our three-species models are well approximated by a single-species model in the sense of [@burgess].
However:
*1)* Because the interaction terms in \eqref{eq:3d model} are about an order of magnitude smaller than $r_X$ and $r_Y$, Models 2-4 are ‘‘close’’ in parameter-space to a single-species model.
*2)* The fact that for Model 2 all strategies match in performance suggests that \eqref{eq:3d model} might be well-approximated by a single species model.
This in turn suggests that the reason that DRL outperforms both single-species strategies for Models 3 and 4 is not due a lack of a single-species approximation for either $X$ or $Y$, but due to the complexity of having two harvested species.
Moreover, non-stationarity exacerbated the advantage of DRL over CEsc and CMort, which can be expected as non-stationary parameters have been observed to introduce bias into single-species reference point estimation [@burgess].

@@

## Future directions

There are a number of interesting directions that we intend to explore in future work.

First, it is to benchmark our results on increasingly more compex and realistic fishery models.
This would include non-stationarities in the dynamical parameters to accurately reflect the effects of climate change on the ecosystem.
This added complexity would likely pose a computational challenge---in future work we will likely need to test several different DRL training algorithms (see e.g. [@rl-intro] for a non-exhaustive list), and it is very likely that hyperparameter tuning will need to be performed.
Moreover, it may be that larger neural networks than the one we used in this research will be needed for the policy function.
This all will mean that considerable technical work will be needed in order to make this next step computationally feasible (e.g. we might need to make more extensive use of GPUs and parallelization).

@@

# Appendix: Results for stationary models

@@

# Appendix: PPO policy function for non-stationary model

@@

# Appendix: Gaussian process interpolation
